Collaboration in Ad Hoc Teamwork: \\ Ambiguous Tasks, Roles, and Communication\vspace{1em}

### 2015-06-25 reviews/rebuttal_ijcai2015.md

Thanks to all reviewers for your comments.

## Reviewer 1:

This paper presents a significant extension to the ad hoc teamwork
problem. Ad hoc team research focuses on the problem where all agents
know the domain except the ad hoc agent, and those agents cannot be
modified. Therefore, we cannot change the other agents to have them
tell the new agents what to do. An example of this scenario is when a
new robot is deployed to a location where there are already older
robots working. The ad hoc agent can only influence them using the
actions in the world and thus cannot directly exchange the required
information. Also, the communication protocol itself is unknown.  Our
main contribution is the formulation of this complex problem in a way
that can be addressed by an efficient online method based on a
Bayesian filter.

## Reviewer 2:

We thank the reviewer for all the valuable insights that we can use
for a longer version of the paper and in future work.  It is true that
the difficulty in this domain arises from the agents' limited
observability, as our agent can quickly interpret the task and its
desired role when it can observe all the agents.  Our algorithm is
designed for scenarios with partial observability, where the ad hoc
agent must interpret the ambiguous messages of its teammates.

## Review 3:

An example of adhoc teamwork arises when a new robot is deployed to a
location where there are already older robots working.  The main
challenge is that solving for all of the variables simultaneously is
computationally expensive and if care is not taken it will be easy to
get stuck in local minima.

The difficulty of the problem does not come from delayed feedback;
even with instantaneous rewards the problem would exist. The
challenge is that the agent does not observe the reward and cannot
estimate it directly because it does not know which task needs to be
done.

We used value iteration to compute the policies.

We agree that theoretical results that would be interesting include
the impact of using approximate spaces of domains and using
approximate filtering methods to reduce computational complexity.
Future work will consider how to remove the assumption that predators
see each other.

### 2015-11-29 paper/adhoccom.tex
update readme for beta release
updating the paper given reivew, removing claim on algortihm (there were almost none)

\documentclass{article}

\usepackage{ijcai15}
\usepackage{times}

@@

@@

\title{Collaboration in Ad Hoc Teamwork: \\
Ambiguous Tasks, Roles, and Communication}

\author{Anonymous authors for blind review}

% \author{
% Jonathan Grizou \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ jonathan.grizou@inria.fr
% \And
% Samuel Barrett \\ Kiva Systems \\ North Reading, MA 01864 USA \\ basamuel@kivasystems.com
% \AND
% Peter Stone \\ Dept. of Computer Science \\ The Univ. of Texas at Austin \\ Austin, TX 78712 USA \\ pstone@cs.utexas.edu
% \And
% Manuel Lopes \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ manuel.lopes@inria.fr
% }

@@

\begin{abstract}
Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents. Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. In this paper, we consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. Our approach assumes a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that an ad hoc agent can integrate itself into a team by quickly identifying the task, selecting its role on the team, and estimating the communication protocol being used; thus barely impacting the overall team performance compared to a pre-coordinated agent.
\end{abstract}

@@

% \section*{Acknowledgments}
%
% Work partially supported by INRIA, Conseil
% R\'{e}gional d'Aquitaine, the ERC grant EXPLORERS 24007, and a INRIA Explorer fellowship.

\clearpage

\bibliographystyle{named}
\bibliography{ijcai15}

>>>>

\documentclass[a4paper]{article}

\usepackage{conference}
\usepackage{times}

@@

\usepackage{url}

@@

\title{Collaboration in Ad Hoc Teamwork: \\
Ambiguous Tasks, Roles, and Communication\vspace{1em}}

% \author{Anonymous authors for blind review}
\setlength\titlebox{3.8in}

\author{
Jonathan Grizou \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ jonathan.grizou@inria.fr
\And
Samuel Barrett \\ Kiva Systems \\ North Reading, MA 01864 USA \\ basamuel@kivasystems.com
\AND
Peter Stone \\ Dept. of Computer Science \\ The Univ. of Texas at Austin \\ Austin, TX 78712 USA \\ pstone@cs.utexas.edu
\And
Manuel Lopes \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ manuel.lopes@inria.fr
}

@@

\begin{abstract}
Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents. Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. In this paper, we consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that an ad hoc agent can integrate itself into a team by quickly identifying the task, selecting its role on the team, and estimating the communication protocol being used; thus barely impacting the overall team performance compared to a pre-coordinated agent.
\end{abstract}

@@

\section*{Open Science}

The code developed for this work is available online at \url{https://github.com/jgrizou/adhoc_com}.

\section*{Acknowledgments}

Work partially supported by INRIA, Conseil R\'{e}gional d'Aquitaine, the ERC grant EXPLORERS 24007, and a INRIA Explorer fellowship.

% \clearpage

\bibliographystyle{named}
\bibliography{adhoccom}

### 2015-11-29 paper/introduction.tex
updating the paper given reivew, removing claim on algortihm (there were almost none)

There are many situations where an international effort is needed to address a particular well focused problem, e.g. rescue efforts in a natural disaster area. In this setting, multiple robots might be available, but they do not have the same software and hardware, and they will not communicate and coordinate using standard protocols. Nevertheless, they should coordinate to achieve a common goal, even if team coordination strategies cannot be pre-defined.

@@

In this work, we focus on the ad hoc team challenge (\cite{stone2010ad}). We imagine a team of specialized agents that work to achieve a specific task and coordinate using a specific language. We introduce a new algorithm allowing an ad hoc agent to join and collaborate with such a team. The ad hoc agent must take its role in the team and must therefore identify the three following components:
%
\begin{itemize}
\item the \textbf{task} the team is trying to solve, so as to help the team achieve it;
\item the \textbf{role} of each agent, so as to replace the missing specialized agent;
\item the \textbf{communication protocol} used by the team, so as to be informed of important facts concerning the task.
\end{itemize}
%
We present empirical results in the pursuit domain showing that an ad hoc agent can efficiently replace any member of such a team. For this purpose, we assume the ad hoc agent has access to a set of hypotheses about the possible tasks, team configurations, and communication systems. Given this information it is possible to infer which hypothesis is the most likely given the observations of other agents' movements and communications. We show that the default team performance is quickly recovered after our ad hoc agent is included and further show that our algorithm is robust to noise.

%In the following sections, we formally define the problem and describe how an ad hoc agent can infer which hypothesis is correct one select its actions. We then describe the pursuit domain used for testing our algorithm, as well as the possible tasks, team configurations, and communication systems. In Section \ref{sec:results}, we compare the default team performance with the one including the ad hoc agent. Finally, we propose a number of directions for future work and highlight related work from other fields of research.

@@

This paper differs from previous work in that the ad hoc agent is not informed of the task to be achieved and does not initially understand the communication of the other agents. This work shares similarities with previous work in that our team includes specialized agents, as in \cite{genter2011role}, and we will assume a finite set of possible teammates and domain configurations, as in \cite{genter2011role,barrett2011empirical,barrett2014communicating}.

@@

A similar problem, where both the task and the communication are unknown, has already been investigated in human-machine interaction \cite{grizou2014interactive}. They consider only two agents, a teacher and a learner. The learning agent has access to even less information than the scenario presented here -- it could not observe the actions of the other agent and could only act on its own towards the success of the task. However, the learner could observe non-symbolic communication signals from the other agent, whose meaning, i.e. its language, was initially unknown. Bringing their work to the multi-agent problem is not straightforward and would require additional developments not considered in this work.

>>>>

There are many situations where an international effort is needed to address a particular well focused problem, e.g. rescue efforts in a natural disaster area. In this setting, multiple robots might be available, but they are unlikely to have the same software and hardware, and they will not communicate using standard protocols. Nevertheless, they should be able to coordinate to achieve a common goal, even if team coordination strategies cannot be pre-defined.

@@

In this work, we focus on the ad hoc team challenge (\cite{stone2010ad}). We imagine a team of specialized agents that work to achieve a specific task and coordinate using a specific language. We replace one of these agents with an ad hoc agent that should learn to collaborate with the team. The ad hoc agent must take its role in the team and must therefore identify all of the three following components:
%
\begin{itemize}
\item the \textbf{task} the team is trying to solve, so as to help the team achieve it;
\item the \textbf{role} of each agent, so as to replace the missing specialized agent;
\item the \textbf{communication protocol} used by the team, so as to be informed of important facts concerning the task.
\end{itemize}
%
We present empirical results in the pursuit domain showing that an ad hoc agent can efficiently replace any member of such a team. For this purpose, we assume the ad hoc agent has access to a set of hypotheses about the possible tasks, team configurations, and communication systems. Given this information it is possible to infer which hypothesis is the most likely given the observations of other agents' movements and communications. We show that the default team performance is quickly recovered after our ad hoc agent is included. We further introduce partial observability and noise on the agent's actions and communication.

% In the following sections, we define the problem and describe our approach. We then describe the pursuit domain used for testing the algorithm, as well as the possible tasks, team configurations, and communication systems. In Section \ref{sec:results}, we compare the default team performance with the one including the ad hoc agent. Finally, we propose a number of directions for future work and highlight related work from other fields of research.

@@

This paper differs from previous work in that \textit{the ad hoc agent is not informed of the task to be achieved and does not initially understand the communication of the other agents}. Our main contribution is the formulation of this complex problem in a way that can be addressed by an online method based on a Bayesian filter. This work shares similarities with previous work in that our team includes specialized agents, as in \cite{genter2011role}, and we will assume a finite set of possible teammates and domain configurations, as in \cite{genter2011role,barrett2011empirical,barrett2014communicating}.

@@

A similar problem, where both the task and the communication are unknown, has already been investigated in human-machine interaction \cite{grizou2014interactive}. They consider only two agents, a teacher and a learner. The learning agent has access to less information than the scenario presented here -- it could not observe the actions of the other agent and could only act on its own towards the success of the task. However, the learner could observe non-symbolic communication signals from the other agent, whose meaning, i.e. its language, was initially unknown. It is yet unclear if their approach can be adapted to multi-agent scenarios.

% Bringing their work to the multi-agent problem is not straightforward and would require additional developments not considered in this work.

### 2015-11-29 paper/problem.tex
updating the paper given reivew, removing claim on algortihm (there were almost none)

Observing the behavior of all other agents is a valuable source of information. Given a hypothesis domain $d_h$, we can compute the probability of the next agent state $S'$ given the current agent state $S$. For each hypothesis, we create a Bayes' filter that accumulates the probability of each domain conditionally on the observation of the agent movements.
To do so, we must estimate the probability that each agent selected each available action. We then estimate the probability of the observed state given all possible combinations of agents' actions and the environment dynamics:
% We denote that the computational cost of such update rule rises quickly with the number of agents.
%
\begin{eqnarray}
p(d_h|S',S) &\propto& p(S'|S,d_h) p(d_h)
\label{eq:obsupdate}
\end{eqnarray}
%
with
%
\begin{eqnarray}
p(S'|S,d_h) \prod_i \sum_j p(s'_{b_i}|s_{b_i},u_j,\env_h) p(u_j|s_{b_i}, S, d_h)
\label{eq:obsupdateexpanded}
\end{eqnarray}
%
where $\env_h$ is the environment in $d_h$ which includes the state transition model. And $p(u_j|s_{b_i}, S, d_h)$ is the model of agent $b_i$ action selection, which is based on all the components of $d_h$ and the current state of the domain $S$. Given the agents' roles, they actions are independent. % But this might not always be the case, as exemplified in the pursuit domain example of section~\ref{sec:domain}. To accommodate with thisThe equation above considers the case of full knowledge of the states. As this might not always be true, the update rule should also account for partial knowledge of states.
The equation becomes:
%
\begin{eqnarray}
p(d_h) &=& \sum_{S'} \sum_S p(d_h|S',S) p(S') p(S)
\label{eq:obsupdatestateprob}
\end{eqnarray}
%
which can be expanded as for Equation~\ref{eq:obsupdateexpanded}. For example, this uncertainty about the state of the world can be explained by partial observability from the ad hoc agent. Communication between agents can help refine its estimate of the current state.

@@

We now consider the action selection method for the ad hoc agent. Previous work considered ad hoc agents that solve the optimal teammate problem. Such agents know the model of their teammates and select their next action in order to improve maximally the team performance -- often resulting in better performance that the initial team \cite{barrett2011empirical}. The aim of this study is different; we want to advocate the fact that the ad hoc agent can work under less assumptions than before and be able to estimate more information about the new team. Hence, we isolate our algorithm from the planning aspects and test whether it can select between candidate domain. Therefore, in this work, the ad hoc agent simply tries to replace a missing agent in the team. To this end, the ad hoc agent will weight the policies of the agent it replaces for each domain hypotheses $d_h$ by the probability currently assigned to this configuration $p(d_h)$.
%
\begin{eqnarray}
p(u_a|M, S,d_h) = \sum_h p(u_a|M, S, d_h) p(d_h)
\label{eq:adhocpolicy}
\end{eqnarray}
%
With this planning strategy, once the correct hypothesis is identified, the ad hoc agent will mimic the default behavior of the agent it replaces.% Nothing limits us to the use of this policy and more advanced planning strategies -- which have already been shown to be beneficial in this kind of scenario \cite{barrett2011empirical} -- could be considered.

>>>>

Observing the behavior of all other agents is a valuable source of information. Given a hypothesis domain $d_h$, we can compute the probability of the next agent state $S'$ given the current agent state $S$. For each hypothesis, we create a Bayes' filter that accumulates the probability of each domain conditionally on the observation of the agent movements.
To do so, we must estimate the probability that each agent selected each available action. We then estimate the probability of the observed state given all possible combinations of agents' actions and the environment dynamics:
% We denote that the computational cost of such update rule rises quickly with the number of agents.
%
\begin{eqnarray}
p(d_h|S',S) &\propto& p(S'|S,d_h) p(d_h)
\label{eq:obsupdate}
\end{eqnarray}
%
with
%
\begin{eqnarray}
p(S'|S,d_h) = \prod_i \sum_j p(s'_{b_i}|s_{b_i},u_j,\env_h) p(u_j|s_{b_i}, S, d_h)
\label{eq:obsupdateexpanded}
\end{eqnarray}
%
where $\env_h$ is the environment in $d_h$ which includes the state transition model. And $p(u_j|s_{b_i}, S, d_h)$ is the model of agent $b_i$ action selection, which is based on all the components of $d_h$ and the current state of the domain $S$. Given the agents' roles, they actions are independent. % But this might not always be the case, as exemplified in the pursuit domain example of section~\ref{sec:domain}. To accommodate with thisThe equation above considers the case of full knowledge of the states. As this might not always be true, the update rule should also account for partial knowledge of states.
The equation becomes:
%
\begin{eqnarray}
p(d_h) &=& \sum_{S'} \sum_S p(d_h|S',S) p(S') p(S)
\label{eq:obsupdatestateprob}
\end{eqnarray}
%
which can be expanded as for Equation~\ref{eq:obsupdateexpanded}. For example, this uncertainty about the state of the world can be explained by partial observability from the ad hoc agent. Communication between agents can help refine its estimate of the current state.

@@

We now consider the action selection method for the ad hoc agent. Previous work considered ad hoc agents that solve the optimal teammate problem. Such agents know the model of their teammates and select their next action in order to improve maximally the team performance -- often resulting in better performance that the initial team \cite{barrett2011empirical}. The aim of this study is different; we want to advocate the fact that the ad hoc agent can work under less assumptions than before and be able to estimate more information about the new team. Hence, we isolate our algorithm from the planning aspects and test whether it can select between candidate domain. Therefore, in this work, the ad hoc agent simply tries to replace a missing agent in the team. To this end, the ad hoc agent will weight the policies for each domain hypotheses $d_h$ by the probability currently assigned to this configuration $p(d_h)$.
%
\begin{eqnarray}
p(u_a|M, S,d_h) = \sum_h p(u_a|M, S, d_h) p(d_h)
\label{eq:adhocpolicy}
\end{eqnarray}
%
With this planning strategy, once the correct hypothesis is identified, the ad hoc agent will mimic the default behavior of the agent it replaces. But the agent is likely to make sensible decision earlier as unrelevant hypotheses are discarded.

% Nothing limits us to the use of this policy and more advanced planning strategies -- which have already been shown to be beneficial in this kind of scenario \cite{barrett2011empirical} -- could be considered.
